---
title: "Chess Project"
author: "Scott Karr"
date: "February 28, 2016"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

0. Load Packages (these may not all be used)
```{r load-pkgs, eval=TRUE}
library(stringr)
library(XML)
library(RCurl)
library(bitops)
library(tau)
library(plyr)
library(dplyr)
```

referenced for initial load . . .
<http://stackoverflow.com/questions/21114598/importing-a-text-file-into-r>

1. Load Data Frame from website
```{r load-text-by-line, eval=TRUE}
theUrl <- "/Users/scottkarr/IS607Spring2016/project1/tournamentinfo.txt"
l <- readLines(theUrl)
```
remove unnecessary lines
```{r remove-header, eval=TRUE}
l <- grep("^\\|?-+\\|?$|^$", l, value = TRUE, invert = TRUE)
```

split
```{r delimit, eval=TRUE}
lsplit <- strsplit(l, "\\s*\\|")
```

set names
```{r name-columns, eval=TRUE}
dat <- setNames(data.frame(do.call(rbind, lsplit[-1])[ ,-1]), paste(lsplit[[1]],lsplit[[2]])[-1])
```

add back last column name
```{r name2-columns, eval=TRUE}
colnames(dat)[10] <- "Pair Num"
```

1st 2 rows were combined for header so remove row 1 which is still left
```{r remove-combined-row, eval=TRUE}
dat <- dat[-c(1), ]
```

convert list to data frame
```{r frame,eval=TRUE}
df1 <- data.frame(dat)
```
subset child and parent recs
```{r subset-into-parent&child, eval=TRUE}
df1[,"IsChildRec"] <- str_detect(df1[,1],"[[:digit:]]{1,}")
df1.Csub <- subset(df1,df1$IsChildRec == TRUE )
df1.Psub <- subset(df1,df1$IsChildRec == FALSE )
colnames(df1.Psub)[1]  <- "Name" 
colnames(df1.Csub)[1]  <- "Name" 
```

build output dataframe
```{r build-output-df, eval=TRUE}
df1.Output <- data.frame(df1.Psub$Pair.Num)
colnames(df1.Output)[1]  <- "ID"
df1.Output["Name"] <- df1.Psub$Name
df1.Output["State"] <- df1.Csub$Pair.Num
df1.Output["Ttl-Pts"] <- df1.Psub$Total..Pts
df1.Output["Pre-Rating"] <- str_trim(str_extract(str_trim(df1.Csub$Name), "[:blank:][:digit:]{1,4}"))
df1.Output["Opp1"] <- as.numeric(str_extract(df1.Psub$Round...1, "[:digit:]{1,}$"))
df1.Output["Opp2"] <- as.numeric(str_extract(df1.Psub$Round...2, "[:digit:]{1,}$"))
df1.Output["Opp3"] <- as.numeric(str_extract(df1.Psub$Round...3, "[:digit:]{1,}$"))
df1.Output["Opp4"] <- as.numeric(str_extract(df1.Psub$Round...4, "[:digit:]{1,}$"))
df1.Output["Opp5"] <- as.numeric(str_extract(df1.Psub$Round...5, "[:digit:]{1,}$"))
df1.Output["Opp6"] <- as.numeric(str_extract(df1.Psub$Round...6, "[:digit:]{1,}$"))
df1.Output["Opp7"] <- as.numeric(str_extract(df1.Psub$Round...7, "[:digit:]{1,}$"))
```

Last derived column uses lapply to scan each row and apply  dplyr to filter indexed data point (scores).
While lapply is perhaps more direct then nested loops, I'm uncomfortable with it.  Too much indirection.
Should be able to collapse the hardcoded column indexes as well, but for another time.
```{r apply-calc-avg, eval=TRUE}
df1.Output["AvgOppScore"] <- 
    unlist(
        lapply(
          1:nrow(df1.Output), 
          function(i) {
              mean(
                c(
                      as.numeric(dplyr::filter(df1.Output, as.numeric(ID) == df1.Output[i,6]+1)[5]),
                      as.numeric(dplyr::filter(df1.Output, as.numeric(ID) == df1.Output[i,7]+1)[5]),
                      as.numeric(dplyr::filter(df1.Output, as.numeric(ID) == df1.Output[i,8]+1)[5]),
                      as.numeric(dplyr::filter(df1.Output, as.numeric(ID) == df1.Output[i,9]+1)[5]),
                      as.numeric(dplyr::filter(df1.Output, as.numeric(ID) == df1.Output[i,10]+1)[5]),
                      as.numeric(dplyr::filter(df1.Output, as.numeric(ID) == df1.Output[i,11]+1)[5]),
                      as.numeric(dplyr::filter(df1.Output, as.numeric(ID) == df1.Output[i,12]+1)[5])                                              
                ),
                na.rm = TRUE
              )                  
            }
          )
        )  
```

remove opponent id references 
```{r cleanup, eval=TRUE}
df1.Output <- df1.Output[-c(6:12)]
df1.Output
```

Export to .csv to your current working directory
```{r export-as-csv, eval=TRUE}
write.csv(df1.Output, file = "tournamentinfo.csv")
```